### 2015-05-11 conclusion.tex
some updates to include speedups

In this paper we presented \whoop, a new fully automatic approach that can soundly detect all possible data races in device drivers. Compared to traditional data race detection techniques that are based on happens-before and typically attempt to explore as many thread interleavings as possible (and thus face code coverage and scalability issues), \textsc{Whoop} uses over-approximation, summarization and symbolic pairwise lockset analysis, and thus can scale well. Exploiting the race-freedom guarantees provided by \whoop, we showed that we can achieve a sound partial-order reduction that can significantly accelerate \corral, a state-of-the-art precise bug-finder for concurrent programs. Using the combination of \whoop and \corral, we analyzed \sizeOfBenchmarks drivers from the 4.0 Linux kernel distribution, showing that our approach can analyze programs 1.5--20$\times$ faster than standalone \corral.